### 2025-03-10 paper/paper.qmd
change paper structure and update references/bibliography

---
title: "Propaganda Films and National Identity: Animated Films in USSR and China"
shorttitle: "Propaganda Films and National Identity"
author:
  - name: Lexun Yu
    email: lx.yu@mail.utoronto.ca
    affiliations:
    - name: University of Toronto
course: "BMS401: Animated Film in Europe"
professor: Laura Pontieri
duedate: March 20, 2025
bibliography: references.bib
documentmode: stu
draftall: true
floatsintext: true
author-note:
  correspondence-note: "Code, drafts, citations, referenced sources, and figures are available at: [https://github.com/yulexun/propaganda-film-analysis](https://github.com/yulexun/propaganda-film-analysis)"
format:
  apaquarto-pdf: default
---

The Cold War is a geopolitical and ideological war that defines global politics between the United States and the Soviet Union emerged after World War II from 1947 to 1991. The conflict between the two countries extends beyond the military advancement, reaching political, economical, and cultural areas [@riabovImagesUrbanSpace2021, pp. 123]. Animated films, as one of the cultural products, often include the use of propaganda as a tool to shape public opinion and assert ideological dominances [@pontieriSovietAnimationThaw2012, pp. 6]. The USSR and its allies promoted communism as a superior alternative to capitalism, which is dominant in the United States. Both the USSR and the newly established People’s Republic of China needs cultural products to prove the rightness of their ideological system. During the 1950s and 1960s, both the USSR and China utilize cinema as a medium for propaganda to gain public support for their own views on socialism. In this essay, we compare six propaganda animated films (three from the USSR and three from PRC) from the two countries in the 1950s and 1960s. We aim to identify the differences between these films and the social-political reasons behind them. First, we introduce the historical background in the post WWII era in both countries. Second, we introduce the six films and the directors. This includes the director's choice of style across different films, depictions of characteristics including movement, background, and sound, and the overall theme, objective, and target audiences. Finally, we connect the characteristics observed in these films and make connections with the general trends in the local animation industry, as well as historical and socio-cultural context. We conclude that ...

# Backgrounds

# Animated Propaganda Films {#sec-films}

The three films we choose from the USSR are Mister Volk directed by @gromovMisterVolk1949, Mister Tvister directed by @karanovichMisterTvister1963, and Millioner directed by @bordzilovskiyMillioner1963. These films are all produced by the state established film studio Soyuzmultfilm.

![Background and characters of MrWolf are detailed and rich.](fig/wolf1.png){#fig-wolf1 width=50%}

# Social and Political Connections {#sec-connections}

Lorem ipsum odor amet, consectetuer adipiscing elit. Primis libero ac lacus class leo interdum. Lacinia nulla rhoncus netus platea conubia porttitor ultrices auctor. Quam consectetur eros accumsan hendrerit feugiat aliquam torquent magnis. Iaculis ridiculus quis enim dignissim aenean metus ipsum hendrerit vitae. Sapien praesent nulla suscipit odio sed non nullam adipiscing. Metus libero turpis potenti varius lobortis mus. Sed consectetur tristique iaculis fringilla; ipsum turpis litora justo quis. Congue neque platea pharetra fermentum metus auctor. Eros penatibus neque ante; in ornare conubia.

Magna urna vestibulum, tortor praesent vel mattis montes. Ultricies sollicitudin interdum habitant magnis vitae suscipit leo montes. Nisl integer pharetra eget facilisi tempus augue dictum penatibus. Torquent proin felis magna etiam; varius nascetur suspendisse suscipit. Inceptos auctor penatibus neque lobortis aptent porttitor. Nibh primis auctor turpis donec semper primis rhoncus. Etiam nulla convallis accumsan felis nunc rhoncus dapibus. Curae dui himenaeos natoque imperdiet adipiscing proin consectetur potenti facilisis. Elementum curae tempus adipiscing risus aptent. Habit·asse porttitor nisl sociosqu; cras nunc facilisis.

Dapibus lacinia nulla eleifend justo luctus vulputate vestibulum taciti massa. Aliquet dictumst maecenas efficitur leo praesent massa sit feugiat curae. Hac lacus enim mattis sapien interdum ut vulputate primis. Ipsum dictum nisi semper; purus facilisi lorem leo ridiculus. Ipsum sem velit nec cursus ornare dis morbi et. Finibus aenean maximus sagittis justo natoque quam velit cras? Malesuada habitant volutpat vulputate dictum massa id egestas fermentum quam. Metus vivamus risus mauris euismod ex natoque ex in pellentesque.

Nullam fringilla donec vel ridiculus id. Condimentum himenaeos iaculis class lacus porttitor ligula massa sociosqu. Dignissim cursus etiam class per justo tortor. Interdum nam nisi ac potenti efficitur eleifend torquent diam. Senectus nulla laoreet ridiculus tellus lobortis; finibus tincidunt ac. Taciti morbi risus facilisi hendrerit adipiscing vitae lacus cubilia ultrices. Feugiat blandit diam lacinia tortor netus condimentum massa imperdiet pretium. Magna quis integer quam netus felis.

Curabitur mi praesent; malesuada parturient vitae adipiscing. Tempus nostra litora aenean pellentesque ante. Tincidunt phasellus per litora porta; tempor orci est. Tellus viverra natoque lobortis mauris praesent. Natoque platea ipsum urna hendrerit primis ac elit, lacus taciti. Tristique tristique enim velit habitant odio diam sollicitudin. Aptent posuere morbi varius ornare dis; ultrices arcu. Torquent montes lacinia vitae libero dictumst.

This another example according to @cowenAnimationIronCurtain2020.

\newpage

# References

::: {#refs}
:::